---
Author: Rajesh Chinnaga
title: "R Notebook"
output: html_notebook
---

---
Assumptions:
1. We cannot have more than one author for a particular book.
2. Author, genre and title are all dependend on the title. 
3. Title of each book is unique.
4. Every book has a unique id.
---

Making a connection with the database:
```{r}
library(RSQLite)
dbpath <- "C:/DBMS/SQL/Databases"  #Path of database
dbfile <- "BookDatebase.db"                 #Name of database

dbcon <- dbConnect(RSQLite::SQLite(), paste0(dbpath, dbfile))
```

Q1] Create a normalized (BCNF) relational schema and visualize the schema in an ERD for the data in the XML file. Include the ERD in your R Notebook.
```{r}
knitr::include_graphics('https://imgur.com/xCBmUbu')
```

Q2] Create a SQLite database that implements the schema, i.e., define the tables with CREATE TABLE. Use SQL chunks in your R Notebook.

```{sql connection=dbcon}
DROP TABLE IF EXISTS Book;
```

```{sql connection=dbcon}
DROP TABLE IF EXISTS Catalog;
```



Creating Book table:
```{sql connection=dbcon}
CREATE TABLE IF NOT EXISTS Book(
  title TEXT NOT NULL,
  author TEXT NOT NULL,
  genre TEXT NOT NULL,
  desc TEXT NOT NULL,
  PRIMARY KEY(title)
)
```


Creating Catalog table:
```{sql connection=dbcon}
CREATE TABLE IF NOT EXISTS Catalog(
  id TEXT NOT NULL,
  title TEXT NOT NULL,
  price TEXT NOT NULL,
  pub_date TEXT NOT NULL,
  edition NUMBER,
  PRIMARY KEY(id, title)
)
```

Q3] Load the XML data from the file into R data frames; you will need to use either node-by-node traversal of the XML tree or a combination of node-by-node traversal with XPath; you likely will not be able to accomplish it with only XPath. Use surrogate keys and/or the ID attributes in the XML to establish relationships.

Loading required libraries:
```{r}
library(DBI)
library(XML)
```

Loading our XML file:
```{r parseXML}
path <- "C:/DBMS"  #Path of the XML file
xmlFile <- "Books-v3.xml"                 #Name of the  XML file

xmlDoc <- xmlParse(validate = F, file = paste(path,xmlFile,sep="/"))  #Parsing the XML file
r <- xmlRoot(xmlDoc)
xmlObj <- xmlDoc
```

Loading XML file into data frames:

Loading data for the book table:
```{r}
xpathEx1 <- "//title"
book <-data.frame(xpathSApply(xmlObj,xpathEx1,xmlValue))
names(book)[1] <- "title"

xpathEx1 <- "//author"
author <-xpathSApply(xmlObj,xpathEx1,xmlValue)
book$author <- author

xpathEx1 <- "//genre"
genre <-xpathSApply(xmlObj,xpathEx1,xmlValue)
book$genre <- genre

xpathEx1 <- "//description"
desc <-xpathSApply(xmlObj,xpathEx1,xmlValue)
book$desc <- desc
book <- book[!duplicated(book), ]
```

Loading data for the Catalog Table
```{r}
xpathEx2 <- "//book[@id]"
catalog <-data.frame(xpathSApply(xmlObj,xpathEx2,xmlGetAttr,"id"))
names(catalog)[1] <- "id"

xpathEx2 <- "//title"
title <-xpathSApply(xmlObj,xpathEx2,xmlValue)
catalog$title <- title

xpathEx2 <- "//price"
price <-xpathSApply(xmlObj,xpathEx2,xmlValue)
catalog$price <- price

xpathEx2 <- "//publish_date"
pub_date <-xpathSApply(xmlObj,xpathEx2,xmlValue)
catalog$pub_date <- pub_date

xpathEx2 <- "//edition"
e1 <-xpathSApply(xmlObj,xpathEx2,xmlValue)
e2 <-xpathSApply(xmlObj,"//book[edition][@id]", xmlGetAttr, "id")
catalog$edition <-NA
catalog[catalog$id==e2, ]$edition<-e1
```

Q4] Transform data types as necessary and then save the data frames to the appropriate tables in the database using a strategy of your choice (e.g., dbWriteTable()).

```{r}
dbWriteTable(dbcon, "Book",book, append=T,row.names=F)
dbWriteTable(dbcon, "Catalog",catalog, append=T,row.names=F)
```

Q5] Once the data from the XML is in the database, build SQL chunks for the following queries:

A. 
```{sql connection=dbcon}
SELECT COUNT(genre) as "Genre_Count "
  FROM (SELECT COUNT(*) as genre
  FROM Book
  GROUP BY genre
  HAVING count(*)>=3
  )

```


B. 

```{sql connection=dbcon}
SELECT strftime('%Y', MAX(pub_date)) as "Most_recent_year"
FROM Book INNER JOIN Catalog ON Book.title = Catalog.title
```


C.

```{sql connection=dbcon}
SELECT genre,COUNT(*) as "no_of_books", avg(price) as "avg_price"
FROM Book INNER JOIN Catalog ON Book.title=Catalog.title
GROUP BY genre
```


D.
```{sql connection=dbcon}
SELECT Catalog.title, author, price
FROM Book INNER JOIN Catalog ON Book.title=Catalog.title
WHERE 
price < 0.8*( SELECT avg(price) FROM Catalog)
AND 
price > 1.2*( SELECT avg(price) FROM Catalog)
```


